---
title: "Analysis of Teachers' Recruitment in Saudi Arabia Between 1437 and 1440"
subtitle: "Data from Saudi Open Data Office"
author: "Musab Isah"
date: "6 March 2021"
output: html_document
---

Group Number:

Group Members' Names (IDs):

1.AHMED ADNAN MOHAMMED ABUALHNOUD (2200000997)-- Team Leader
2.-- Member 1
3.-- Member 2
4.-- Member 3



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries 

Ensure you have the `tidyverse` and `readxl` packages installed by running the 2 lines of code below. Both the team leader and members should run the code below in the console of rstudio.

```{r libraries, echo=FALSE, message = FALSE}
Sys.setlocale("LC_CTYPE", "arabic" )
library(tidyverse)
library(readxl)
if(!require(ggpubr)){
   install.packages("ggpubr")
   library("ggpubr")
}
```


We then need to load the data from the MS Excel (.xlsx) file, named teachers_data.xlsx, to the R environment for processing.

*Note: Only Team Leader should run the code below*

```{r load_data, message = FALSE}
tdata <- read_excel("teachers_data.xlsx")
```

# Wherever you see the text '#Add some code', it means you should delete the comment and add your code for the task.

# 2 Marks will be deducted for every deadline missed.

## Task 1
\textcolor{red}{(8 points)}

```{r tidy_data}
colnames(tdata) <- c("ID", "Administrative_Region", "Levels",

                     "Authority", "Education_Office", "Sex",

                     "School_Type", "Education_Type",

                     "School_System", "Saudis", "Expats",
 
                     "Total_Numb_Teachers", "Year")
```

Team Leader should *Knit, commit, and push changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

\textcolor{red}{DEADLINE: END OF WEEK 9}

## Task2
\textcolor{red}{(8 points)}

Next Team Member should *Pull the changes made by Team Leader before proceeding.*

(a) What are the different administrative_regions (Al-Mandaqah Al-idariyyah) found in the data?

```{r admin_regions}
tdata %>%
  distinct(Administrative_Region)
```

Comments about your answer:There are 13 different administrative regions

(b) Based on your knowledge of the administrative regions in Saudi Arabia, is there any region that has not recruited any teacher during the period under review (1437 - 1440 Hijri)? List the Saudi regions shown in the data.

```{r not_recruited_admin_regions}
unique(tdata$Administrative_Region)
```

Comments about your answer:All 13 region have teachers recruited and have teachers 

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 10}

## Task3
\textcolor{red}{(20 points)}

The Next Team Member should *Pull the changes made by Member 1 before proceeding.*

(a) Saudi Arabian education is divided into different stages/levels (Al-Marhala), list the different levels showed in the data in descDEADLINE: ENDing order of frequency. (4 points)

```{r study_levels}
tdata %>%
  count(Levels) %>%
  arrange(desc(n))
```

Comments about your answer: most of the recruitments were done(المرحلة الابتدائية)by primary school  because the level appears 58,471 times in the dataset.


The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

The Next Team member should *Pull the changes made by Team Leader before proceeding.*

(b) Plot the graph of region versus number of **Saudi** teachers recruited over the period. (4 points)

```{r plot_admin_regions_saudi}
saudi_teachers <- tdata %>%
  group_by(Administrative_Region) %>%
  summarise(Saudis = sum(Saudis))
view(saudi_teachers)

ggplot(saudi_teachers, aes(x = Administrative_Region, y = Saudis)) +

 geom_point() +

 theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Comments about your answer: we could see from the graph that the Riyadh region  recorded the most recruitment of Saudis followed closely by the Mecca region. North Hudud recruited the least number of teachers.

(c) Plot the graph of region versus number of **Non-Saudi** teachers recruited over the period. (4 points)

```{r plot_admin_regions_non-saudi}
non_saudi_teachers <- tdata %>%
  group_by(Administrative_Region) %>%
  summarise(Expats = sum(Expats))
view(non_saudi_teachers)

ggplot(non_saudi_teachers, aes(x = Administrative_Region, y =Expats)) +

 geom_point() +

 theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Comments about your answer:Comments about your answer:The Riyadh region recruited more expatriate teachers than other regions. And Al Baha recruited the least number of teachers.

(d) Plot the graph of region versus number of **all teachers** recruited over the period. (4 points)

```{r plot_admin_regions_all}
all_teachers <- tdata %>%
  group_by(Administrative_Region) %>%
  summarise(Total = sum(Total_Numb_Teachers))
ggplot(all_teachers, aes(x = Administrative_Region, y = Total)) +

 geom_point() +

 theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Comments about your answer: the graph show all teachers.

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 12}

## Task 4
\textcolor{red}{(12 points)}

The Next Team member should *Pull the changes made previously before proceeding.*

(a) How many teachers are recruited in boys' and girls' schools?  (4 points)

```{r no_boys_girls}
tdata %>%
  group_by(Sex) %>%
  summarise(Total_recruit_teachers= sum(Total_Numb_Teachers))
```

Comments about your answer:The number of teachers are recruited in boys les than girls .

(b) Between boys and girls schools, which recruit more teachers? (4 points)

Comments about your answer: Girls' schools recruited more teachers, 1,166,895, than boys' schools, 954,028.

(c) What is the correlation between recruitment at boys schools and recruitment of girls schools? (4 points)

```{r plot_correlation_boys_girls}

pivoted <- tdata %>%
 group_by(Sex) %>%
 mutate(row = row_number()) %>%
 pivot_wider(
  names_from = Sex,
  values_from = Total_Numb_Teachers)
colnames(pivoted)[13] <- "Boys"
colnames(pivoted)[14] <- "Girls"
pivoted <- pivoted %>%
 select(row, Boys, Girls)
my_fun <- function(x) x[!is.na(x)]
pivoted <- pivoted %>%
 group_by(row) %>%
 summarise_all(funs(my_fun))
ggscatter(pivoted, x = "Boys", y = "Girls",

     add = "reg.line", conf.int = TRUE,

     cor.coef = TRUE, cor.method = "pearson",

     xlab = "Boys", ylab = "Girls")

```

Comments about your answer:

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 13}

## Task 5
\textcolor{red}{(12 points)}

The Next Team member should *Pull the changes made previously before proceeding.*

(a) How many teachers are recruited each year, 1437, 1438, 1439, and 1440? 

```{r teachers_per_year}

no_1437 <- tdata %>%

 filter(Year == "1437") %>%

 summarise(sum(Total_Numb_Teachers))

cat(paste0("1437 Recruits: `", no_1437, "`\n"))

no_1438 <- tdata %>%

 filter(Year == "1438") %>%

 summarise(sum(Total_Numb_Teachers))

cat(paste0("1438 Recruits: `", no_1438, "`\n"))

no_1439 <- tdata %>%

 filter(Year == "1439") %>%

 summarise(sum(Total_Numb_Teachers))

cat(paste0("1439 Recruits: `", no_1439, "`\n"))

no_1440 <- tdata %>%

 filter(Year == "1440") %>%

 summarise(sum(Total_Numb_Teachers))

cat(paste0("1440 Recruits: `", no_1440, "`\n"))
```

Comments about your answer:

(b) Which region recruited the most teachers in each of the years under review? (4 points)

```{r teachers_per_region}
most_teachers <- tdata %>%
 group_by(Administrative_Region, Year) %>%
 summarise(Total = sum(Total_Numb_Teachers)) %>%
 arrange(Year, Total)
most_teachers <- most_teachers[c(13, 26, 39, 52),]
most_teachers

```

Comments about your answer:

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 14}

## If you are here, then congratulations!! You have completed the DSC 200 course project.